Race-based disparities in allocation of academic disciplinary actions are associated

### 2018-03-22 written_docs/draft.Rmd
edit the plotting code

plot.dat3$bias <- 'Explicit'
plot.dat3$metric <- fct_recode(plot.dat3$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
plot.dat3_all <- rbind(plot.dat3_all, plot.dat3)
ggplot(plot.dat3_all, aes(x=metric, y=est, group=bias, color=bias, shape=bias)) +
  geom_point(position=position_dodge(width=.5), size=2) +
  geom_errorbar(aes(ymin=lower, ymax=upper), 
                width=.5, position=position_dodge(width=.5)) +
  theme_classic() +
  coord_flip() +
  geom_hline(yintercept=0) +
  ylab('Black-White Difference in log-odds slope') +
  theme(axis.title.y = element_blank(),
        legend.title = element_blank())
```

@@

```{r detail-risks, fig.cap="Association between bias and relative risk ratio for black students to white students. Line is the mean of the posterior. Bands indicate 95% uncertainty intervals. Points represent counties, whose sizes are scaled to the number of schools in that county. Note that the y axis is cut at 8 for legibility despite data extending beyond."}
plot.dat <- data.frame(group=NA, weighted_bias=NA, est=NA, 
                       lower=NA, upper=NA, bias=NA, metric=NA)
plot.dat2 <- data.frame(weighted_bias=NA, est=NA, 
                        lower=NA, upper=NA, bias=NA, metric=NA)
raw_or <- data.frame(county_id=NA, nschools=NA, bias=NA, 
                     value=NA, black=NA, white=NA, odds_r=NA, metric=NA)
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

@@

  posterior_distribution %>%
    select(group, weighted_bias, sample, value, index) %>%
    group_by(group, weighted_bias) %>%
    summarise(est = mean(value),
              lower = quantile(value, .025),
              upper = quantile(value, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Implicit',
           metric = i) %>%
    rbind(plot.dat) -> plot.dat

  posterior_distribution %>%
    select(group, weighted_bias, sample, value) %>%
    spread(group, value) %>%
    mutate(ratio = black/white) %>%
    group_by(weighted_bias) %>%
    summarise(est = mean(ratio),
              lower = quantile(ratio, .025),
              upper = quantile(ratio, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Implicit',
           metric=i)  %>%
    rbind(plot.dat2) -> plot.dat2

@@

  posterior_distribution %>%
    select(group, weighted_explicit_diff, sample, value, index) %>%
    mutate(weighted_bias = weighted_explicit_diff) %>%
    group_by(group, weighted_bias) %>%
    summarise(est = mean(value),
              lower = quantile(value, .025),
              upper = quantile(value, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Explicit',
           metric=i) %>%
    rbind(plot.dat) -> plot.dat

  posterior_distribution %>%
    select(group, weighted_explicit_diff, sample, value) %>%
    spread(group, value) %>%
    mutate(ratio = black/white,
           weighted_bias = weighted_explicit_diff) %>%
    group_by(weighted_bias) %>%
    summarise(est = mean(ratio),
              lower = quantile(ratio, .025),
              upper = quantile(ratio, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Explicit',
           metric=i) %>%
    rbind(plot.dat2) -> plot.dat2

@@

plot.dat$metric <- fct_recode(plot.dat$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')

plot.dat2$metric <- fct_recode(plot.dat2$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')

raw_or$metric <- fct_recode(raw_or$metric, 
                            'Out-of-School Suspension' = 'oos_susp',
                            'In-School Suspension' = 'inschool_susp',
                            'School-Related Arrest' = 'in_school_arrest',
                            'Law Enf. Referral' = 'law_enforcement',
                            'Expulsion' = 'expulsion_combined')

plot.dat %>%
  filter(!is.na(weighted_bias)) %>%
  ggplot(aes(x=weighted_bias, y=est, group=group, linetype=group)) +
    geom_line(aes(color=group)) +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.15) +
    theme_classic() +
    ylab('Probability of \n disciplinary action') +
    xlab('Standardized Bias') + 
    facet_grid(metric~bias, scales='free') +
    xlim(-3, 3) +
    theme(legend.title = element_blank()) -> p1

plot.dat2 %>% 
  filter(!is.na(weighted_bias)) %>%
  ggplot(aes(x=weighted_bias, y=est)) +
    geom_point(data=raw_or, aes(x=value, y=odds_r, size=nschools), alpha=.05) +
    geom_line() +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.25) +
    theme_classic() +
    scale_size_continuous(name='Number\nof schools') + 
    #scale_color_manual(values=c('grey', 'darkorange'), guide=F) + 
    scale_alpha_manual(values=c(.2, 1), guide=F) +
    ylim(0,8) +
    xlim(-3, 3) + 
    xlab('Standardized Bias') +
    ylab('Relative \n Risk Ratio') +
    geom_hline(yintercept=1) +
    facet_grid(metric~bias) -> p2

@@

p1

>>>>

plot.dat3$bias <- 'Explicit'
plot.dat3$metric <- fct_recode(plot.dat3$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
plot.dat3_all <- rbind(plot.dat3_all, plot.dat3)
p <- ggplot(plot.dat3_all, aes(x=metric, y=est, group=bias, color=bias, shape=bias)) +
  geom_point(position=position_dodge(width=.5), size=3) +
  geom_errorbar(aes(ymin=lower, ymax=upper), 
                width=.5, position=position_dodge(width=.5)) +
  theme_classic() +
  coord_flip() +
  geom_hline(yintercept=0) +
  ylab('Black-White Difference in log-odds slope') +
  theme(axis.title.y = element_blank(),
        legend.title = element_blank(),
        axis.text = element_text(size=12),
        axis.title.x = element_text(size=14, face='bold'),
        legend.text = element_text(size=12))
ggsave('/Users/travis/Documents/gits/educational_disparities/figs/slope_ests.jpeg', p, width = 11, height = 6)
```

@@

```{r detail-risks, fig.cap="Association between bias and relative risk ratio for black students to white students. Line is the mean of the posterior. Bands indicate 95% uncertainty intervals. Points represent counties, whose sizes are scaled to the number of schools in that county. Note that the y axis is cut at 8 for legibility despite data extending beyond."}
plot.dat <- data.frame(group=NA, weighted_bias=NA, est=NA, lower=NA, 
                       upper=NA, bias=NA, metric=NA, p_neg_black=NA, p_neg_white=NA)
plot.dat2 <- data.frame(weighted_bias=NA, est=NA, lower=NA, 
                        upper=NA, bias=NA, metric=NA, p_neg=NA)
raw_or <- data.frame(county_id=NA, nschools=NA, bias=NA, 
                     value=NA, black=NA, white=NA, odds_r=NA, metric=NA)
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

@@

  p_neg_blackimp <- sum(p_cons[3,]>0)/4000
  p_neg_blackexp <- sum(p_cons[4,]>0)/4000
  p_neg_whiteimp <- sum((p_cons[3,]+p_cons[16,])>0)/4000
  p_neg_whiteexp <- sum((p_cons[4,]+p_cons[17,])>0)/4000

  posterior_distribution %>%
    select(group, weighted_bias, sample, value, index) %>%
    group_by(group, weighted_bias) %>%
    summarise(est = mean(value),
              lower = quantile(value, .025),
              upper = quantile(value, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Implicit',
           metric = i,
           p_neg_black=p_neg_blackimp,
           p_neg_white=p_neg_whiteimp) %>%
    rbind(plot.dat) -> plot.dat

  p_neg <- sum(p_cons[16,]<0)/4000

  posterior_distribution %>%
    select(group, weighted_bias, sample, value) %>%
    spread(group, value) %>%
    mutate(ratio = black/white) %>%
    group_by(weighted_bias) %>%
    summarise(est = mean(ratio),
              lower = quantile(ratio, .025),
              upper = quantile(ratio, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Implicit',
           metric=i,
           p_neg = p_neg)  %>%
    rbind(plot.dat2) -> plot.dat2

@@

  posterior_distribution %>%
    select(group, weighted_explicit_diff, sample, value, index) %>%
    mutate(weighted_bias = weighted_explicit_diff) %>%
    group_by(group, weighted_bias) %>%
    summarise(est = mean(value),
              lower = quantile(value, .025),
              upper = quantile(value, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Explicit',
           metric=i,
           p_neg_black=p_neg_blackexp,
           p_neg_white=p_neg_whiteexp) %>%
    rbind(plot.dat) -> plot.dat

  p_neg <- sum(p_cons[17,]<0)/4000

  posterior_distribution %>%
    select(group, weighted_explicit_diff, sample, value) %>%
    spread(group, value) %>%
    mutate(ratio = black/white,
           weighted_bias = weighted_explicit_diff) %>%
    group_by(weighted_bias) %>%
    summarise(est = mean(ratio),
              lower = quantile(ratio, .025),
              upper = quantile(ratio, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Explicit',
           metric=i,
           p_neg=p_neg) %>%
    rbind(plot.dat2) -> plot.dat2

@@

plot.dat$metric <- fct_recode(plot.dat$metric, 
                               'Out-of-School \n Suspension' = 'oos_susp',
                               'In-School \n Suspension' = 'inschool_susp',
                               'School-Related \n Arrest' = 'in_school_arrest',
                               'Law Enf. \n Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')

plot.dat2$metric <- fct_recode(plot.dat2$metric, 
                               'Out-of-School \n Suspension' = 'oos_susp',
                               'In-School \n Suspension' = 'inschool_susp',
                               'School-Related \n Arrest' = 'in_school_arrest',
                               'Law Enf. \n Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')

raw_or$metric <- fct_recode(raw_or$metric, 
                            'Out-of-School \n Suspension' = 'oos_susp',
                            'In-School \n Suspension' = 'inschool_susp',
                            'School-Related \n Arrest' = 'in_school_arrest',
                            'Law Enf. \n Referral' = 'law_enforcement',
                            'Expulsion' = 'expulsion_combined')

plot.dat <- plot.dat[!is.na(plot.dat$weighted_bias),]

plot.dat %>%
  filter(weighted_bias==2.5) %>%
  mutate(text_y = est,
         text_x = 2.75) %>%
  select(group, text_y, text_x, metric, p_neg_black, p_neg_white, bias) %>%
  distinct() %>%
  gather(group1, p_neg, p_neg_black, p_neg_white) %>%
  mutate(group1 = str_sub(group1, 7, -1)) %>%
  filter(group==group1) %>%
  select(-group1) -> text_data

text_data$text_comp <- '=='
text_data$text_comp[text_data$p_neg>.99] <- '>'
text_data$text_comp[text_data$p_neg<.01] <- '<'
text_data$p_neg[text_data$p_neg>.99] <- .99
text_data$p_neg[text_data$p_neg<.01] <- .01

plot.dat %>%
  filter(!is.na(weighted_bias)) %>%
  ggplot(aes(x=weighted_bias, y=est, group=group, linetype=group)) +
    geom_line(aes(color=group)) +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.15) +
    theme_classic() +
    geom_text(data=text_data, parse=T, size=2.5, 
              aes(x=2.8, y=text_y, vjust=0,
                  label=paste0('p[pos]', text_comp, round(p_neg,2)))) +
    ylab('Probability of \n disciplinary action') +
    xlab('Standardized Bias') + 
    facet_grid(metric~bias, scales='free') +
    xlim(-2.75, 3.25) +
    theme(legend.title = element_blank()) -> p1

plot.dat2$text_x <- -2.75
plot.dat2$text_y <- 5
plot.dat2$text_comp <- '=='
plot.dat2$text_comp[plot.dat2$p_neg>.99] <- '>'
plot.dat2 %>%
  select(text_x, text_y, text_comp, p_neg, metric, bias) %>%
  distinct() %>%
  mutate(p_neg = round(p_neg, 2)) %>%
  filter(!is.na(p_neg)) -> text_data
text_data$p_neg[text_data$p_neg==1] <- .99

plot.dat2 %>% 
  filter(!is.na(weighted_bias)) %>%
  ggplot(aes(x=weighted_bias, y=est)) +
    geom_point(data=raw_or, aes(x=value, y=odds_r, size=nschools), alpha=.1) +
    geom_line() +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.25) +
    theme_classic() +
    scale_size_continuous(name='Number\nof schools') + 
    geom_text(data=text_data, parse=T, size=3,
              aes(x=-2.75, y=4, label=paste0('p[pos]', text_comp, round(p_neg,2)))) +
    #scale_color_manual(values=c('grey', 'darkorange'), guide=F) + 
    scale_alpha_manual(values=c(.2, 1), guide=F) +
    ylim(0,6) +
    xlim(-3, 3) + 
    xlab('Standardized Bias') +
    ylab('Relative \n Risk Ratio') +
    geom_hline(yintercept=1) +
    facet_grid(metric~bias) +
    theme(legend.position = 'top',
          strip.text.y=element_text(size=6)) -> p2
ggsave('/Users/travis/Documents/gits/educational_disparities/figs/rel_risk.jpeg', p2, width = 11, height = 6)

@@

ggsave('/Users/travis/Documents/gits/educational_disparities/figs/rates.jpeg', p1, width = 11, height = 6)
